Well-quasi-orderings on word languages

### 2025-10-13 src/prelims.tex

\paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
\AP
An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\intro*\omegaOrd$ to
denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
usual ordering. We also use $\intro*\omegaOne$ to denote the first \emph{uncountable}
ordinal.
We only assume superficial familiarity with ordinal arithmetic, and
refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
II]{KRIVINE71} for a detailed introduction to this domain.
Given a tree $T$
whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
say that $\alpha_T$ is the \emph{rank} of $T$.

\AP
Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
decreasing sequences, and the tree of \kl{antichains}. The rank of these
respective trees are called respectively the \intro{maximal order type} of $X$
written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
$X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. These three
parameters are called the \intro{ordinal invariants} of a
\kl{well-quasi-ordered} set $X$. 
As an example, for $(\Nat, \leq)$, all bad sequences are descending and there are no antichains. Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is 1. \todo{Please check this!}
We refer to the survey of \cite{DZSCSC20} for
a detail discussion on these concepts and their computation on specific classes
of well-quasi-ordered sets.

>>>>

\paragraph*{Ordinal Invariants.} \label{ordinal-invariants:subsec}
\AP
An \intro{ordinal} is a \kl{well-founded} \kl{totally ordered}
set. We use $\alpha, \beta, \gamma$ to denote ordinals, and use $\intro*\omegaOrd$ to
denote the first infinite \kl{ordinal}, i.e., the set of natural numbers with the
usual ordering. We also use $\intro*\omegaOne$ to denote the first \emph{uncountable}
ordinal.
We only assume superficial familiarity with ordinal arithmetic, and
refer to the books of Kunen \cite{KUNEN80} and Krivine~\cite[Chapter
II]{KRIVINE71} for a detailed introduction to this domain.
Given a tree $T$
whose branches are all finite we can define an \kl{ordinal} $\alpha_T$ inductively
as follows: if $T$ is a leaf then $\alpha_T = 0$, if $T$ has children
$\seqof{T_i}$ then $\alpha_T = \sup \setof{\alpha_{T_i} + 1}{i \in \Nat}$. We
say that $\alpha_T$ is the \intro{rank} of $T$.

\AP
Let $(X, \leq)$ be a \kl{well-quasi-ordered} set. One can define three
well-founded trees from $X$: the tree of \kl{bad sequences}, the tree of
decreasing sequences, and the tree of \kl{antichains}. The \kl{rank} of these
respective trees are called respectively the \intro{maximal order type} of $X$
written $\intro*{\oType{X}}$ \cite{dejongh77}, the \intro{ordinal height} of
$X$ written $\intro*{\oHeight{X}}$ \cite{schmidt81}, and the \intro{ordinal
width} of $X$ written $\intro*{\oWidth{X}}$ \cite{kriz90b}. These three
parameters are called the \intro{ordinal invariants} of a
\kl{well-quasi-ordered} set $X$. 
As an example, for $(\Nat, \leq)$, all bad sequences are descending and there are no antichains. Therefore the maximal order type and the ordinal height of $\Nat$ are $\omega$ and the ordinal width is 1. \todo{Please check this!}
We refer to the survey of \cite{DZSCSC20} for
a detail discussion on these concepts and their computation on specific classes
of well-quasi-ordered sets.
